Making Sequential Programs Deterministic

### 2019-04-02 experiments.tex
edits to randoop case study

\section{deflaker case study}
\begin{table}[]
    \begin{tabular}{|l|l|}
        \hline
        Benchmark, issue & found by det checker? \\ \hline
        achilles, [2] issue \#309 &  Race condition\\ \hline
        achilles, [3] issue \#310 & Same fix for \#309 and \#310 \\ \hline
        checkstyle [4] issue \#4664& det checker on test files \\ \hline
        jackrabbit-oak [6] \#6512&  Unresolved on issue tracker \\ \hline
        jackrabbit-oak [7] issue \#6524&  Dead link, didn't find 6524 in issue tracker \\ \hline
        nutz [10] issue \#1283 &  Removed tests (Printing (HTTP's) Response.getstatus())\\ \hline
        togglz [19] issue \#223 & build issue \\ \hline
        togglz [20] issue \#240 &  \\ \hline
    \end{tabular}
\end{table}

>>>>

\section{deflaker case study}
\begin{table}[]
    \begin{tabular}{|l|p{5cm}|}
        \hline
        Benchmark, issue & found by det checker? \\ \hline
        achilles, [2] issue \#309 &  Race condition\\ \hline
        achilles, [3] issue \#310 & Same fix for \#309 and \#310 \\ \hline
        checkstyle [4] issue \#4664&test (added to test cases in det checker) \\ \hline
        jackrabbit-oak [6] \#6512&  Unresolved on issue tracker (Race condition issue) \\ \hline
        jackrabbit-oak [7] issue \#6524&  Dead link, didn't find 6524 in issue tracker \\ \hline
        nutz [10] issue \#1283 &  Removed tests (Printing (HTTP's) Response.getstatus()) has Chinese characters,
        compilation with det checker stuck!\\ \hline
        togglz [19] issue \#223 & build issue \\ \hline
        togglz [20] issue \#240 &  build issue\\ \hline
    \end{tabular}
\end{table}

### 2019-04-02 randoop-case-study.tex
edits to randoop case study

The developers took extensive action to detect and mitigate the effect of, nondeterminism.
They used Docker images to run tests, to avoid system dependencies such as
a different JDK having a different number of classes or methods.
They wrote tests with relaxed oracles (assertions) that permit multiple
possible answers --- for example, in code coverage of generated tests.
They used linters such as Error Prone, which warns if toString is used on
array objects, which don't override Object.toString and therefore print a
hash code which may vary from run to run.
They used libraries that make hash codes deterministic, by giving each
object of a type a unique ID that counts up from 1, rather than using a
memory address as \<Object.hashCode> does.
They wrote preprocess output and logs to make nondeterministic ones easier
to compare, such as by removing or canonicalizing hash codes, dates, and
other nondeterministing output.
Each of these efforts wase helpful, but they were not enough to solve the problem.

In July 2017, the Randoop developers spent two weeks of full-time work to
eliminate unintentional nondeterministic behavior in Randoop.
Their methodology was to repeatedly
 run Randoop with verbose logging enabled,
 look for differences in logging output,
 find the root cause of nondeterminism,
 and eliminate it.
The most common causes were toString() routines and iteration order of sets and maps.
Some of the nondeterminism was in libraries, such as the JDK\@.

The developers chose not to make every Map a LinkedHashMap, because that
would have increased memory and CPU costs.  They chose not to make every
order-nondeterminism List a Set, for similar reasons:  deduplication was
not always needed and would have increased costs.

@@

\TheDeterminismChecker found three previously-unknown nondeterminism bugs in Randoop.
The Randoop developers accepted our bug reports and committed fixes to the repository.

@@

\item 
 Diagnostic output that printed a hash code for brevity, but should have had deterministic output\commit{661a4970}.  By contrast to the other two bugs, this one is not serious.

\item
Dependency on the CLASSPATH environment variable in preference to the
classpath passed on the command line\commit{330e3c56}.
This could have led to incorrect behavior if a user set the environment variable.
The developers changed Randoop to not read the environment variable.
\end{itemize}

We reported another suspicous case of order-nondeterminism, in the
\<findOverridden> method.  The Randoop developers explained it was
acceptable, after tracing the flow through the program.  (One said, ``This
looks OK to me \ldots\ it took me a little while to decide that.'')  The fact
depended on subtle, undocumented invariants about Randoop that we had not
been able to reverse-engineer.

>>>>

The developers took extensive action to detect and mitigate the effect of nondeterminism.
They used Docker images to run tests, to avoid system dependencies such as
a different JDK having a different number of classes or methods.
They wrote tests with relaxed oracles (assertions) that permit multiple
possible answers --- for example, in code coverage of generated tests.
They used linters such as Error Prone, which warns if \<toString> is used on
array objects, which don't override \<Object.toString> and therefore print a
hash code which may vary from run to run.
They used libraries that make hash codes deterministic, by giving each
object of a type a unique ID that counts up from 1, rather than using a
memory address as \<Object.hashCode> does.
They wrote preprocess output and logs to make nondeterministic ones easier
to compare, such as by removing or canonicalizing hash codes, dates, and
other nondeterministing output.
Each of these efforts was helpful, but they were not enough to solve the problem.

In July 2017, the Randoop developers spent two weeks of full-time work to
eliminate unintentional nondeterministic behavior in Randoop.
Their methodology was to repeatedly
 run Randoop with verbose logging enabled,
 look for differences in logging output,
 find the root cause of nondeterminism,
 and eliminate it.
The most common causes were \<toString()> routines and iteration order of sets and maps.
Some of the nondeterminism was in libraries, such as the JDK\@.

The developers chose not to make every \<Map> a \<LinkedHashMap>, because that
would have increased memory and CPU costs.  They chose not to make every
order-nondeterministic \<List> a \<Set>, for similar reasons:  deduplication was
not always needed and would have increased costs.

@@

\TheDeterminismChecker found three previously-unknown nondeterminism bugs in Randoop.
The Randoop developers accepted our bug reports and committed fixes to the repository. A summary
of these bugs follows:

@@

\item
Dependency on the CLASSPATH environment variable in preference to the
classpath passed on the command line\commit{330e3c56}.
This could have led to incorrect behavior if a user set the environment variable.
The developers changed Randoop to not read the environment variable.

\item 
Diagnostic output that printed a hash code for brevity, but should have had deterministic output\commit{661a4970}.  By contrast to the other two bugs, this one is not serious.
\end{itemize}

We reported another suspicious case of order-nondeterminism, in the
\<findOverridden> method.  The Randoop developers explained it was
acceptable, after tracing the flow through the program.  (One said, ``This
looks OK to me \ldots\ it took me a little while to decide that.'')  The fact
depended on subtle, undocumented invariants about Randoop that we had not
been able to reverse-engineer.
